### 2022-03-23 paper/localpsf.tex
moved discretization to intro

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise. For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$. I.e., 
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}

The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
\begin{equation}
\label{eq:defn_of_Akerpcmesh}
\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
\end{equation}
Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
\begin{equation}
\label{eq:kernel_representation_pcmesh}
(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
\end{equation}
where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
\end{equation}
and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the sparse finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
\end{equation*}
Our matrix representation, $\mathbf{A}$, of $\Aop$ is therefore given by
\begin{equation}
\label{eq:Amatpc_defn}
\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
\end{equation}
Note that $\mathbf{A}$ is not exactly the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$. Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.

Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are not easily accessible. Directly computing and storing these matrices is computationally intractible. The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\mathbf{A}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.

@@

\section{Efficient matrix-free approximation of operators via impulse response interpolation: main ideas}
\label{sec:prerequisites}

@@

\section{Efficient matrix-free approximation of operators via impulse response interpolation: construction of the H-matrix}
\label{sec:method}

In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.

>>>>

In computations, functions are discretized and replaced by finite dimensional vectors, and operators mapping between infinite dimensional spaces are replaced by operators mapping between finite dimensional spaces. In this paper we discretize using finite elements. With minor modifications, our method could be used with other discretization schemes such as finite differences or finite volumes.

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise. Further, let $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denote the sparse finite element mass matrix which has entries
\begin{equation*}
\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
\end{equation*}
For $u_h \in V_h \subset L^2(\Omega)$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}
Linear functionals $\rho_h \in V_h' \supset L^2(\Omega)'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries $\left(\boldsymbol{\rho}_h\right)_i = \rho_h(\febasis_i)$ for $i=1,\dots,N$. Here $\mathbb{R}^N_\mathbf{M}$ is the vector space $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$. Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\mu_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\mu}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\mu_h(u) = \boldsymbol{\mu}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\mu_h, \nu_h \in V_h'$. The Riesz representation $\mu_h^* \in V_h$ of $\mu \in V_h'$ becomes $\boldsymbol{\mu}^* = \mathbf{M}^{-1} \boldsymbol{\mu}$. After Galerkin approximation, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ becomes an operator $\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has matrix entries $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$. Notice that $\mathbf{A}$ is dense, and computing one matrix entry $\mathbf{A}_{ij}$ requires an action of $\Aop$, which is already computationally expensive (though we get the entire $j$th column as a side effect of computing $\mathbf{A}_{ij}$). Hence methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$. Computing and storing a matrix representation of $\mathbf{A}$ is out of the question.

If functions in $V_h$ are continuous at $x$, then the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta} \in V_h'$ with entries $\boldsymbol{\delta}_i = \febasis_i(x)$, $i=1,\dots,N$. The Riesz representative of the discrete delta distribution is $\boldsymbol{\delta}^* = \mathbf{M}^{-1} \boldsymbol{\delta}_x$, and the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes $\boldsymbol{\impulseresponse}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$.

\subsubsection{Discretized integral kernel}

The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
\begin{equation}
\label{eq:defn_of_Akerpcmesh}
\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
\end{equation}
Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations, yields the following approximation:
\begin{equation}
\label{eq:kernel_representation_pcmesh}
(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u},
\end{equation}
where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
\end{equation}
Based on \eqref{eq:kernel_representation_pcmesh}, we define the following matrix approximation, $\widetilde{\mathbf{A}} \approx \mathbf{A}$ as follows:
\begin{equation}
\label{eq:Amatpc_defn}
\widetilde{\mathbf{A}} = \bm{M} \mathbf{\Phi} \bm{M}.
\end{equation}
Note that $\widetilde{\mathbf{A}}$ differs slightly from $\mathbf{A}$ (the Galerkin projection of $\Aop$ onto $V_h$) because of the approximation $\Phi^h \approx \Phi$. 
%Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$. 
%Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, 
The matrices $\widetilde{\mathbf{A}}$ and $\mathbf{\Phi}$ are, of course, not easily accessible.  
%The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\widetilde{\mathbf{A}}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.

@@

\subsubsection{Lumped mass matrix and discrete kernel non-negativity}

\section{Key innovations}
\label{sec:prerequisites}

@@

\section{Method}
\label{sec:method}

%In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.
